\chapter{Models with Time--Varying Dynamics for Pandemic A(H1N1) Influenza in Finland}
\label{chap:lna_extensions}

\section{Overview}
\label{sec:lna_extensions_overview}
To this point, we have largely worked with stochastic epidemic models (SEMs) where the transmission dynamics of an outbreak are taken to be constant over time. This might be only mildly unreasonable for short outbreaks in closed, relatively ``well--mixed" populations, and is often an attractive modeling choice as SEMs with static dynamics are easier to interpret and fit. Incidence data typically arise in settings where the outbreak milieu can change due to environmental changes, heterogeneity in the contact structures of various subpopulations as they are exposed, or behavioral responses as people become aware of an outbreak (or complacent about the extent to which it is under control). Furthermore, we are often interested in understanding the effects of interventions that are time--varying in their implementation or action, such as vaccination campaigns, on the transmission dynamics of an outbreak. Thus, it is important, if not necessary, that we allow the time--varying aspects of an outbreak to be flexibly expressed in the model.

In this chapter, we will use SEMs where the force of infection is time--varying to model the spread of pandemic A(H1N1) influenza in Finland using surveillance data. Our goals will be to quantify the transmission dynamics of the outbreak, to estimate the true incidence, and to understand what effect a national vaccination campaign had in mitigating the outbreak severity. We will fit several models using the linear noise approximation (LNA) framework developed in Chapter \ref{chap:lna_for_sems}. The models range in complexity and flexibility, and consequently in the computational and statistical challenges we will have to overcome.

\subsection{Pandemic A(H1N1) Influenza in Finland}
\label{subsec:flu_description}

\subsection{The Importance of Allowing for Time--Varying Dynamics}
\label{subsec:tparam_motivation}

\section{Models with Time--Varying Dynamics}
\label{sec:flu_tparam_models}

\subsection{Modeling Time--Varying Force of Infection with a Fixed Functional Form}
\label{subsec:flu_quadfoi}

\subsection{Flexible Models for the Force of Infection with Gaussian Markov Random Fields}
\label{subsec:flu_gmrf}

\subsection{More models}

\section{Discussion}